\textbf{Course: Física Teórica I (CDP7600)} \\[0.2cm] {\Large \textbf{Project}: \hspace{0.2cm} \textit{Fibration building blocks of information-processing networks} }

### 2019-11-08 Manuscript/main.tex
finishing manuscript

The formal definition of an input-tree is explained in detail at \cite{fibration2019}, \cite{Boldi2006} and \cite{Boldi2002}. Here, I give just a direct and intuitive example through the examples showed at figures \ref{fig:ex}. To fully characterize the information flow directions in a directed network, we can define for each node in the network an input-tree containing all the information pathways that ends at the given node. Thus, if two different nodes have isomorphic input-trees, which means that they receives equivalent information from the network, they synchronize their dynamical states, exhibiting a correlated behavior for the network. Even more, if a group of nodes have isomorphic input-trees, then a graph fibration transformation reduces these nodes from network $G$ in one single node for a base network $G'$. This way, a fibration reduces the original network in a base network where each node of $G'$ represents a fiber block of $G$. In figure \ref{fig:ex} we can see a directed network formed just by three nodes. For each one, they differs only by the patterns of directed links, from which information can flow.
\begin{figure}[h]
	\centering
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex1.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex2.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex3.png}}}
	\caption{First trial network example. Containing a total of $N = 17$ nodes and $M = 22$ arcs, the network has two weakly connected components, defined as the connected networks when we treat all the directed arcs as undirected edges. For each non-trivial fiber I give the same color for the nodes belonging to it.}
	\label{fig:ex}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex1input.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex2input.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex3input.png}}}
	\caption{First trial network example. Containing a total of $N = 17$ nodes and $M = 22$ arcs, the network has two weakly connected components, defined as the connected networks when we treat all the directed arcs as undirected edges. For each non-trivial fiber I give the same color for the nodes belonging to it.}
	\label{fig:inputex}
\end{figure}

@@

The algorithm \ref{alg:branchratio} assumes that the given node $v$ is part of a strongly connected component, even for the trivial case containing only $v$ connected to itself, because this warrant that $n \neq 0$. This can be easily verified implementing a Depth-First Search or Breadth-First Search on the network and its tranpose starting at node $v$ to find its strongly connected component, which takes linear time with the size of the strongly connected component. Then, for components containing only fiber nodes the algorithm obtain the correct integer $n$ for small $\delta$ limit tree depth, running linearly with $|E^{-1}(\{V_{SCC}\}) \in \Lambda|$. For component containing outsider nodes, then we define $\Lambda$ containing the nodes in the shortest loop \ref{fig:branch} and apply the above algorithm. Of course, for fibonacci loops, determining the size $s$ of the shortest loop already gives indirectly the information $\varphi_s$ needed, being necessary only to define previously a table with the relations $s \rightarrow \varphi_s$.

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.32]{Figures/input.png}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \subfloat[]{{\includegraphics[scale=0.22]{Figures/result1-label.png}}}
    \qquad
    \begin{tabular}[b]{cc}\hline
      Fiber & $| n, \ell \rangle$  \\ \hline
      1 & $| 0, 2 \rangle$ \\
      2 & $| 0, 1 \rangle$ \\
      3 & $| 0, 1 \rangle$ \\ 
      4 & $| 0, 1 \rangle$ \\
      5 & $| 0, 2 \rangle$ \\ \hline
	\end{tabular}
	\captionlistentry[table]{A table beside a figure}
    %\captionsetup{labelformat=andtable}
    \caption{A table beside a figure}
\end{figure}

@@

%\begin{figure}[H]
%	\centering
%	\subfloat[]{{\includegraphics[scale=0.25]{Figures/result2.png}}}
%	\qquad
%	\subfloat[]{{\includegraphics[scale=0.25]{Figures/result2-label.png}}}
%	\caption{First trial network example. Containing a total of $N = 17$ nodes and $M = 22$ arcs, the network has two weakly connected %components, defined as the connected networks when we treat all the directed arcs as undirected edges. For each non-trivial fiber I %give the same color for the nodes belonging to it.}
%	\label{fig:result2}
%\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.275]{Figures/result2.png}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.275]{Figures/result2-label.png}
\end{figure}

\begin{table}[H]
\centering
\begin{tabular}[b]{cc}\hline
	Fiber & $| n, \ell \rangle$  \\ \hline
	1 & $| 0, 1 \rangle$ \\
	2 & $| 0, 1 \rangle$ \\
	3 & $| 0, 1 \rangle$ \\ 
	4 & $| 0, 1 \rangle$ \\
	5 & $| 0, 2 \rangle$ \\
	6 & $| 0, 2 \rangle$ \\ \hline
  \end{tabular}
  \caption{}
  \label{tab:r2}
\end{table}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.27]{Figures/result3.png}
	\caption{}
	\label{fig:result3}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.27]{Figures/result3-1.png}
	\caption{}
	\label{fig:result3-1}
\end{figure}

The connected component showed at the figure \ref{fig:result3} is a subset representation of the nodes of the regulatory network of the \textit{Escherichia Coli} bacteria and the resulted fibers shows just a portion of the whole fibers pattern. In the table X I provide the classification for each fibration building block obtained for the whole bacteria network, including, of course, the fibers presented in the figure \ref{fig:result3-1}. In total, I have found $92$ non-trivial fiber groups containing a total of $416$ nodes, obtaining a characteristic fiber statistics for the \textit{Escherichia Coli} network data. To guarantee the correctness of the fiber distribution found I calculated, after the refinement algorithm, the information received for each node inside the fibers and compares if each one process equivalent information. The result of this calculation is stored in the data file given at \url{https://github.com/higorsmonteiro/fiberblocks/blob/master/Data/fiberVerification.dat}, where for each node in a non-trivial fiber block I give the list of the nodes that sends information messages to the current node along with their fiber indexes and type of link values. Thus, the results of this report show an equivalent fiber distribution over the network comparing with results exhibited at \cite{fibration2019} for the same regulatory network data. Even though the table IV an V of \cite{fibration2019} show a total number of $91$ fibers and the total number of $414$ of nodes inside non-trivial fibers, the authors \href{https://docs.google.com/spreadsheets/d/1-RG5vR_EGNPqQcnJU8q3ky1OpWi3OjTh5Uo-Xa0PjOc/edit#gid=1973910256}{spreadsheet} reports the total of $416$ nodes, a mismatch which is caused in this report by the presence of a additional fiber of size two and classified as $| 1.6181..., 7 \rangle$ containing the genes \textit{marR} e \textit{marA}. This way, the number of nodes in fibers and the genes inside each one are consistent with the results showed in \cite{fibration2019}.

@@

\begin{table}[H]
	\centering
	\begin{tabular}[b]{lc}\hline
		Structure Fiber & Amount in E. Coli  \\ \hline
		$| n = 0, \ell = 1 \rangle$ & 45 \\
		$| n = 0, \ell = 2 \rangle$ & 13 \\
		$| n = 0, \ell = 3 \rangle$ & 3 \\ 
		$| n = 1, \ell = 0 \rangle$ & 13 \\
		$| n = 1, \ell = 1 \rangle$ & 8 \\
		$| n = 2, \ell = 1 \rangle$ & 3 \\
		$| n = 2, \ell = 1 \rangle$ & 1 \\ 
		$| n = 1.3802..., \ell = 1 \rangle$ & 1 \\ 
		$| n = 1.4655..., \ell = 1 \rangle$ & 1 \\  
		$| n = 1.6181..., \ell = 2 \rangle$ & 1 \\ 
		$| n = 1.6181..., \ell = 7 \rangle$ & 1 \\ 
		\hline
		\textbf{Total number of building blocks} & 92 \\
		\hline
		\textbf{Total number of nodes inside fibers} & 416 \\
		\hline
	  \end{tabular}
	  \caption{tab:fiberstats}
	  \label{tab:fiberstats}
\end{table}

>>>>

The formal definition of an input-tree is explained in detail at \cite{fibration2019}, \cite{Boldi2006} and \cite{Boldi2002}. Here, I give just a direct and intuitive example through the examples showed at figures \ref{fig:ex}. To fully characterize the information flow directions in a directed network, we can define for each node in the network an input-tree containing all the information pathways that ends at the given node. Thus, if two different nodes have isomorphic input-trees, which means that they receives equivalent information from the network, they synchronize their dynamical states, exhibiting a correlated behavior for the network. Even more, if a group of nodes have isomorphic input-trees, then a graph fibration transformation reduces these nodes from network $G$ in one single node for a base network $G'$. This way, a fibration reduces the original network in a base network where each node of $G'$ represents a fiber block of $G$. In figure \ref{fig:ex} we can see a directed network formed just by three nodes. For each one, they differs only by the patterns of directed links, from which information can flow.
\begin{figure}[h]
	\centering
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex1.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex2.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.4]{Figures/ex3.png}}}
	\caption{The patterns of directed connections can influence the collective dynamical states of the network nodes through synchronization. Considering only the topology, we say that two nodes in a information-processing network synchronize their states if they process equivalent information. Since $Y$ and $Z$ receive same information from the same nodes in (a) e (c), they synchronize, different from the case (b), where all nodes have different dynamical states.}
	\label{fig:ex}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex1input.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex2input.png}}}
	\qquad
	\subfloat[]{{\includegraphics[scale=0.32]{Figures/ex3input.png}}}
	\caption{Nodes in a network that have isomorphic input-tree share the same dynamical states. (a) e (c) shows symmetrical input-tree for nodes $Y$ e $Z$, where due to the addition of the autoregulation loop in (c), their input-trees become infinite.}
	\label{fig:inputex}
\end{figure}

@@

The algorithm \ref{alg:branchratio} assumes that the given node $v$ is part of a strongly connected component, even for the trivial case containing only $v$ connected to itself, because this warrants that $n \neq 0$. This can be easily verified implementing a Depth-First Search or Breadth-First Search on the network and its tranpose starting at node $v$ to find its strongly connected component, which takes linear time with the size of the strongly connected component. Then, for components containing only fiber nodes the algorithm obtain the correct integer $n$ for small $\delta$ limit tree depth, running linearly with $|E^{-1}(\{V_{SCC}\}) \in \Lambda|$. For component containing outsider nodes, then we define $\Lambda$ containing the nodes in the shortest loop \ref{fig:branch} and apply the above algorithm. Of course, for fibonacci loops, determining the size $s$ of the shortest loop already gives indirectly the information $\varphi_s$ needed, being necessary only to define previously a table with the relations $s \rightarrow \varphi_s$.

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.32]{Figures/input.png}
	\caption{Example of fibration building block containing more than one cycle. Building the fiber input-tree with all the three nodes from the strongly connected component gives a wrong branching ratio $n$, while considering only the nodes in the shortest cycle gives the right branching ratio $\varphi_2$.}
	\label{fig:branch}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \subfloat[]{{\includegraphics[scale=0.22]{Figures/result1-label.png}}}
    \qquad
    \begin{tabular}[b]{cc}\hline
      Fiber & $| n, \ell \rangle$  \\ \hline
      1 & $| 0, 2 \rangle$ \\
      2 & $| 0, 1 \rangle$ \\
      3 & $| 0, 1 \rangle$ \\ 
      4 & $| 0, 1 \rangle$ \\
      5 & $| 1, 2 \rangle$ \\ \hline
	\end{tabular}
	\captionlistentry[table]{A table beside a figure}
    %\captionsetup{labelformat=andtable}
    \caption{A table beside a figure}
\end{figure}

@@

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.275]{Figures/result2.png}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.275]{Figures/result2-label.png}
\end{figure}

\begin{table}[H]
\centering
\begin{tabular}[b]{cc}\hline
	Fiber & $| n, \ell \rangle$  \\ \hline
	1 & $| 1.6181..., 1 \rangle$ \\
	2 & $| 1, 0 \rangle$ \\
	3 & $| 1.6181..., 2 \rangle$ \\ 
	4 & $| 0, 1 \rangle$ \\
	5 & $| 0, 1 \rangle$ \\
	6 & $| 0, 1 \rangle$ \\ \hline
  \end{tabular}
  \caption{}
  \label{tab:r2}
\end{table}

The connected component showed at the figure \ref{fig:result3} is a subset representation of the nodes of the regulatory network of the \textit{Escherichia Coli} bacteria and the resulted fibers shows just a portion of the whole fibers pattern. In the table X I provide the classification for each fibration building block obtained for the whole bacteria network, including, of course, the fibers presented in the figure \ref{fig:result3-1}. In total, I have found $92$ non-trivial fiber groups containing a total of $416$ nodes, obtaining a characteristic fiber statistics for the \textit{Escherichia Coli} network data. To guarantee the correctness of the fiber distribution found I calculated, after the refinement algorithm, the information received for each node inside the fibers and compares if each one process equivalent information. The result of this calculation is stored in the data file given at \url{https://github.com/higorsmonteiro/fiberblocks/blob/master/Data/Results/DefinitiveFiber.dat}, where for each node in a non-trivial fiber block I give the list of the nodes that sends information messages to the current node along with their fiber indexes and type of link values. Thus, the results of this report show an equivalent fiber distribution over the network comparing with results exhibited at \cite{fibration2019} for the same regulatory network data. Even though the table IV an V of \cite{fibration2019} show a total number of $91$ fibers and the total number of $414$ of nodes inside non-trivial fibers, the authors \href{https://docs.google.com/spreadsheets/d/1-RG5vR_EGNPqQcnJU8q3ky1OpWi3OjTh5Uo-Xa0PjOc/edit#gid=1973910256}{spreadsheet} reports the total of $416$ nodes, a mismatch which is caused in this report by the presence of a additional fiber of size two, classified as $| 1.6181..., 7 \rangle$ containing the genes \textit{marR} e \textit{marA}. This way, the number of nodes in fibers and the genes inside each one are consistent with the results showed in \cite{fibration2019}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.23]{Figures/result3.png}
	\caption{}
	\label{fig:result3}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.23]{Figures/result3-1.png}
	\caption{}
	\label{fig:result3-1}
\end{figure}

@@

\begin{table}[H]
	\centering
	\begin{tabular}[b]{lc}\hline
		Structure Fiber & Amount in E. Coli  \\ \hline
		$| n = 0, \ell = 1 \rangle$ & 45 \\
		$| n = 0, \ell = 2 \rangle$ & 13 \\
		$| n = 0, \ell = 3 \rangle$ & 3 \\ 
		$| n = 1, \ell = 0 \rangle$ & 13 \\
		$| n = 1, \ell = 1 \rangle$ & 8 \\
		$| n = 2, \ell = 1 \rangle$ & 3 \\
		$| n = 2, \ell = 1 \rangle$ & 1 \\ 
		$| n = 1.3802..., \ell = 1 \rangle$ & 1 \\ 
		$| n = 1.4655..., \ell = 1 \rangle$ & 1 \\  
		$| n = 1.6181..., \ell = 2 \rangle$ & 1 \\ 
		$| n = 1.6181..., \ell = 7 \rangle$ & 1 \\ 
		\hline
		\textbf{Total number of building blocks} & 92 \\
		\hline
		\textbf{Total number of nodes inside fibers} & 416 \\
		\hline
	  \end{tabular}
	  \caption{Fibration building blocks statistics of the \textit{Escherechia Coli} genetic regulatory network. The indexation accounts for trivial fibers also.}
	  \label{tab:fiberstats}
\end{table}

\centering
\begin{table}[b]
	\centering
\begin{tabular}{r|r|r|r}
\hline
Fiber ID &  Fiber Size &  $n$ or $\varphi_d$ &  $\ell$ \\ \hline
4 &          12 &             0.0000 &          1 \\        5 &           9 &             1.4655 &          1 \\       10 &           2 &             0.0000 &          1 \\       11 &           2 &             0.0000 &          1 \\       12 &           2 &             0.0000 &          1 \\       13 &           4 &             0.0000 &          1 \\       15 &           2 &             0.0000 &          2 \\       19 &           2 &             0.0000 &          1 \\       20 &           7 &             0.0000 &          1 \\       21 &           2 &             0.0000 &          1 \\       23 &           2 &             1.3802 &          1 \\       25 &           6 &             0.0000 &          1 \\       26 &           2 &             0.0000 &          1 \\       29 &           4 &             0.0000 &          1 \\       30 &           9 &             0.0000 &          1 \\       35 &           2 &             0.0000 &          1 \\       41 &           4 &             0.0000 &          1 \\       43 &          24 &             0.0000 &          1 \\       46 &          11 &             0.0000 &          1 \\       48 &           7 &             0.0000 &          1 \\       49 &           3 &             0.0000 &          1 \\       53 &           2 &             0.0000 &          1 \\       54 &           5 &             0.0000 &          1 \\       55 &           4 &             0.0000 &          1 \\       56 &           3 &             0.0000 &          1 \\       58 &           2 &             0.0000 &          2 \\       66 &           4 &             0.0000 &          2 \\       67 &          13 &             0.0000 &          1 \\       70 &           2 &             0.0000 &          1 \\       71 &          28 &             0.0000 &          1 \\       73 &           4 &             0.0000 &          3 \\       74 &           3 &             0.0000 &          2 \\       75 &           4 &             0.0000 &          1 \\       78 &           4 &             1.0000 &          1 \\       80 &           3 &             0.0000 &          2 \\       83 &           2 &             1.0000 &          1 \\       84 &           2 &             1.6181 &          2 \\       92 &           4 &             0.0000 &          1 \\ 
\end{tabular}
\end{table}

\begin{table}[b]
	\centering
	\begin{tabular}{r|r|r|r}

		Fiber ID &  Fiber Size &  $n$ or $\varphi_d$ &  $\ell$ \\ \hline
		110 &           2 &             0.0000 &          2 \\      114 &           2 &             0.0000 &          2 \\      117 &           2 &             0.0000 &          2 \\      120 &           2 &             0.0000 &          2 \\      128 &           2 &             0.0000 &          3 \\      132 &           3 &             0.0000 &          1 \\      135 &           5 &             0.0000 &          3 \\      137 &           2 &             2.0000 &          1 \\      145 &           2 &             1.0000 &          1 \\      153 &           4 &             1.0000 &          1 \\      181 &           2 &             0.0000 &          1 \\      187 &           8 &             0.0000 &          1 \\      198 &           2 &             1.0000 &          1 \\      208 &           2 &             1.0000 &          2 \\      225 &           2 &             0.0000 &          2 \\      246 &           2 &             1.0000 &          1 \\      249 &           3 &             1.0000 &          1 \\      257 &           3 &             1.0000 &          0 \\      269 &           2 &             0.0000 &          2 \\      277 &           2 &             2.0000 &          0 \\      278 &           9 &             0.0000 &          1 \\      281 &          11 &             1.0000 &          0 \\      284 &           3 &             0.0000 &          1 \\      299 &          26 &             1.0000 &          0 \\      305 &           3 &             1.0000 &          0 \\      311 &           2 &             0.0000 &          2 \\      312 &           5 &             1.0000 &          0 \\      313 &           2 &             1.0000 &          1 \\      319 &           5 &             0.0000 &          1 \\      322 &          16 &             1.0000 &          0 \\      324 &           3 &             1.0000 &          0 \\      330 &           9 &             1.0000 &          0 \\      331 &           2 &             1.0000 &          0 \\      342 &           3 &             0.0000 &          1 \\      350 &           3 &             0.0000 &          1 \\      353 &           2 &             0.0000 &          1 \\      356 &           2 &             0.0000 &          1 \\
	\end{tabular}
\end{table}

\begin{table}[b]
	\centering
	\begin{tabular}{r|r|r|r}

		Fiber ID &  Fiber Size &  $n$ or $\varphi_d$ &  $\ell$ \\ \hline
				 359 &           2 &             0.0000 &          1 \\      364 &           5 &             0.0000 &          1 \\      373 &          12 &             0.0000 &          1 \\      380 &           2 &             0.0000 &          1 \\      395 &           2 &             0.0000 &          1 \\      433 &           3 &             1.0000 &          2 \\      434 &           2 &             1.0000 &          2 \\      440 &           2 &             0.0000 &          2 \\      449 &           2 &             0.0000 &          1 \\      476 &           2 &             1.6181 &          7 \\      481 &           3 &             0.0000 &          1 \\      495 &           2 &             1.0000 &          0 \\      497 &           2 &             1.0000 &          0 \\      498 &           2 &             1.0000 &          0 \\      500 &           2 &             0.0000 &          1 \\      503 &           2 &             0.0000 &          1 \\      505 &           2 &             1.0000 &          0 \\
				 \hline
	\end{tabular}
	\caption{Size and classification of all non-trivial fibers over the \textit{Escherichia Coli} genetic regulatory network.}
\end{table}
